---
title: "**qcCREST**"
date: 'Last update: `r format(Sys.time(), sep="_", "%Y-%m-%d %H:%M")`'
output: html_document
knit: (function(inputFile, encoding) { 
  rmarkdown::render(inputFile, encoding=encoding, output_file=paste0(substr(inputFile,1,nchar(inputFile)-4), format(Sys.time(), sep="_", "%Y-%m-%d_%H%M"), '.html')) 
  }
  )
---



```{r setup, include=FALSE, echo=F}
knitr::opts_chunk$set(echo=T, warning=F, message=F)

# Load packages ---------------------------
library(dplyr)
library(readxl)
library(writexl)
library(here)


# Read source data ---------------------------
crest_biodata <- read_excel(
  path = 
    # Combining the parent directory, sep, and gsub arguments
    paste0(
      # Creating parent directory:
      paste0(
        "C:/Users", 
        sep="/", 
        # Finds your computer username:
        Sys.info()[6], 
        sep="/",
        # SCA StAD SharePoint website that is synced to your computer:
        "DFO-MPO/PAC-SCA Stock Assessment (STAD) - Terminal CN Run Recon/Communal data"
        ),
      sep="/",
      # Find the file in the directory that starts with "WCVI_Chinook_Run_Reconstruction_Project_Biological_Data_with_FOS" and append to parent directory
      grep("WCVI_Chinook_Run_Reconstruction_Project_Biological_Data_with_FOS",
           list.files(path = paste0("C:/Users", sep="/", Sys.info()[6], sep="/",
                                    "DFO-MPO/PAC-SCA Stock Assessment (STAD) - Terminal CN Run Recon/Communal data" )),
           value=T)
      ),
  sheet="WCVI_Chinook_Run_Rec", 
  guess_max=20000
)


# Helper variables and functions ---------------------------
RR_year <- 2022
"%notin%" <- Negate("%in%")
```

<br>

<br>

Output of script for QCing CREST biodata for terminal run reconstructions. 

Sister file: C:/Users/DAVIDSONKA/DFO-MPO/PAC-SCA Stock Assessment (STAD) - Terminal CN Run Recon/Communal data/qcCREST_date_time.xlsx

<br>

<br>

**1 - Otolith record no sample**

*Otolith box/vial numbers exist but there is 'No Sample'*

```{r}
oto_no_sample <- crest_biodata %>% 
  filter(!is.na(OTOLITH_BOX) & !is.na(OTOLITH_SPECIMEN) & THERMALMARK=="No Sample")
```

`r nrow(oto_no_sample)` cases

<br>




**2 - Scale record no age or explanation**

*Scale book and number but no age, and no explanation given (e.g., resorbed etc)*

```{r}
scale_no_age <- crest_biodata %>% 
  filter(is.na(RESOLVED_AGE) & is.na(PART_AGE_CODE) & !is.na(SCALE_BOOK))
```

`r nrow(scale_no_age)` cases

<br>




**3.1 - CWT head label no result**

*A CWT head label was submitted but the stock ID result is blank*

```{r}
CWT_no_ID <- crest_biodata %>% 
  filter(!is.na(CWT_HEAD_LABEL) & is.na(CWT_RESULT))
```

`r nrow(CWT_no_ID)` cases

<br>




**3.2 - CWT head label no result**

*CWT head label entered as '0'*

```{r}
CWT_zero <- crest_biodata %>% 
  filter(CWT_HEAD_LABEL==0)
```

`r nrow(CWT_zero)` cases

<br>




**4 - Whatman sheet no result**

*Whatman sheet/DNA tracking numbers exist but no GSI result (blank)*

```{r}
whatman_no_sample <- crest_biodata %>% 
  filter(!is.na(SPECIMEN_REFERENCE_DNA_NO) & is.na(DNA_RESULTS_STOCK_1) | DNA_RESULTS_STOCK_1=="NO SAMPLE" ) 
```

`r nrow(whatman_no_sample)` cases

<br>




**5.1 - Resolved_stock_source CWT but DNA disagreement**

*Stock ID assigned by CWT but GSI >=80% disagrees*

```{r}
CWT_DNA_id <- crest_biodata %>% 
  filter(RESOLVED_STOCK_SOURCE=="CWT" & PROB_1>=0.8 & RESOLVED_STOCK_ORIGIN!=REGION_1_NAME)
```

`r nrow(CWT_DNA_id)` cases

<br>




**5.2 - Resolved_stock_source OTO but DNA disagreement**

*Stock ID assigned by otolith thermal mark but GSI >=80% disagrees*

```{r}
oto_DNA_id <- crest_biodata %>% 
  filter(RESOLVED_STOCK_SOURCE=="Otolith Stock" & PROB_1>=0.8 & RESOLVED_STOCK_ORIGIN!=REGION_1_NAME)
```

`r nrow(oto_DNA_id)` cases

<br>




**5.3 - Resolved_stock_source DNA LT 0.8**

*A stock ID assigned by DNA with <80% certainty*

```{r}
DNA_uncert <- crest_biodata %>% 
  filter(RESOLVED_STOCK_SOURCE=="DNA" & PROB_1<0.8)
```

`r nrow(DNA_uncert)` cases

<br>




**5.4 - Hatchery origin blank possible PBT**

*Hatchery origin given as a blank, but some possibility for PBT (PROB=1.00). Note this should be approached with lots of caution and is stock-specific.*

```{r}
PBT_maybe <- crest_biodata %>% 
  filter(is.na(HATCHERY_ORIGIN) & PROB_1==1 & is.na(DNA_STOCK_2))
```

`r nrow(PBT_maybe)` cases

<br>




**5.5 - SUS (assumed) that have ID methods available**

*A CWT head label was submitted but the stock ID result is blank*

```{r}
sus_meta <- crest_biodata %>% 
  filter(RESOLVED_STOCK_ORIGIN=="SUS (assumed)" & 
           !is.na(CWT_RESULT) & CWT_RESULT!="No Tag" & 
           !is.na(THERMALMARK) & !THERMALMARK%in%c("No Sample","Not Marked") &
           !is.na(DNA_RESULTS_STOCK_1) & DNA_RESULTS_STOCK_1!="NO SAMPLE") 
```

`r nrow(sus_meta)` cases

<br>




**6 - Age discrepancy**

*Cases where the RESOLVED_AGE does not match the catch YEAR minus the CWT_BROOD_YEAR*

```{r}
age_discrep <- crest_biodata %>%
  filter((YEAR-CWT_BROOD_YEAR)!=RESOLVED_AGE) 
```

`r nrow(age_discrep)` cases

<br>




**7 - Blank Area_Name records**

*Carry over - not sure if still needed*

```{r}
blank_AreaName <- crest_biodata %>%
  filter(is.na(AREA_NAME))
```

`r nrow(blank_AreaName)` cases

<br>




**8 - Check for non-standard sex designations**

*Sex designation does not fall as M or F - propose changing all other designations to "Unknown"*

```{r}
nonstd_sex <- crest_biodata %>%
  filter(SEX %notin% c("M","F"))
```

`r nrow(nonstd_sex)` cases

<br>









```{r echo=F,warning=F, message=F}
# ======================== CREATE REPORT ========================
# TY Norah Brown for source code/idea
# Summary table --------------------------- 
explore_summary <- data.frame(Issue_ID=character(), Issue=character(), Count=integer(),
                              Definition=character(), 
                              stringsAsFactors=FALSE)

explore_summary <- explore_summary  %>% 
  add_row(Issue_ID="1", Issue="Oto missing", Count=nrow(oto_no_sample), Definition="Cases where there are otolith box/vial numbers but are returning 'No Sample'") %>% 
  add_row(Issue_ID="2", Issue="Scale missing", Count=nrow(scale_no_age), Definition="Scale book and number but no age, and no explanation given (e.g., resorbed etc)") %>% 
  add_row(Issue_ID="3.1", Issue="CWT missing", Count=nrow(CWT_no_ID), Definition="A CWT head label was submitted but the stock ID result is blank") %>% 
  add_row(Issue_ID="3.2", Issue="CWT label 0", Count=nrow(CWT_zero), Definition="CWT head label entered as '0'") %>% 
  add_row(Issue_ID="4", Issue="DNA missing", Count=nrow(whatman_no_sample), Definition="Whatman sheet/DNA tracking numbers exist but no GSI result (blank)") %>% 
  add_row(Issue_ID="5.1", Issue="ID flag CWT-DNA", Count=nrow(CWT_DNA_id), Definition="Stock ID assigned by CWT but GSI >=80% disagrees") %>% 
  add_row(Issue_ID="5.2", Issue="ID flag OTO-DNA", Count=nrow(oto_DNA_id), Definition="Stock ID assigned by otolith thermal mark but GSI >=80% disagrees") %>% 
  add_row(Issue_ID="5.3", Issue="ID flag GSI LT 0.8", Count=nrow(DNA_uncert), Definition="A stock ID assigned by DNA with <80% certainty") %>% 
  add_row(Issue_ID="5.4", Issue="ID flag PBT", Count=nrow(PBT_maybe), Definition="Hatchery origin given as a blank, but some possibility for PBT (PROB=1.00). Note this should be approached with lots of caution and is stock-specific.") %>%
  add_row(Issue_ID="5.5", Issue="ID flag SUS(assumed)", Count=nrow(sus_meta), Definition="Fish identified as Southern US (assumed) but has other information available for stock ID (e.g., GSI)") %>%
  add_row(Issue_ID="6", Issue="Age flag", Count=nrow(age_discrep), Definition="Cases where the RESOLVED_AGE does not match the catch YEAR minus the CWT_BROOD_YEAR") %>%
  add_row(Issue_ID="7", Issue="Area_Name blank", Count=nrow(blank_AreaName), Definition="Carry over - not sure if still needed") %>%
  add_row(Issue_ID="8", Issue="Non-std sex", Count=nrow(nonstd_sex), Definition="Sex designation does not fall as M or F - propose changing all other designations to 'Unknown'")

# Create sheets --------------------------- 
sheet_list<-list(Summary = explore_summary,
                 "1 - Oto missing" = data.frame(oto_no_sample), 
                 "2 - Scale missing" = data.frame(scale_no_age),
                 "3.1 - CWT missing" = data.frame(CWT_no_ID),
                 "3.2 - CWT label 0" = data.frame(CWT_zero), 
                 "4 - DNA missing" = whatman_no_sample,
                 "5.1 - ID flag CWT-DNA" = data.frame(CWT_DNA_id), 
                 "5.2 - ID flag OTO-DNA" = data.frame(oto_DNA_id), 
                 "5.3 - ID flag GSI LT 0.8" = data.frame(DNA_uncert), 
                 "5.4 - ID flag PBT" = data.frame(PBT_maybe),
                 "5.5 - ID flag SUS(assumed)" = data.frame(sus_meta),
                 "6 - Age flag" = data.frame(age_discrep),
                 "7 - Area_Name blank" = data.frame(blank_AreaName),
                 "8 - Non-std sex" = data.frame(nonstd_sex)
)



# ======================== CREATE REPORT ========================

# Detect and delete previous version of qcCREST --------------------------- 
# On SharePoint
unlink(x=paste0(
  paste0(
    "C:/Users", sep="/", Sys.info()[6], sep="/", 
    "DFO-MPO/PAC-SCA Stock Assessment (STAD) - Terminal CN Run Recon/Communal data"
    ),
  sep="/",
  grep("qcCREST",
       list.files(path = paste0("C:/Users", sep="/", Sys.info()[6], sep="/",
                                "DFO-MPO/PAC-SCA Stock Assessment (STAD) - Terminal CN Run Recon/Communal data" )),
       value=T)
  )
)

# On local machine
unlink(x=paste0(
  here("data_prep"),
  grep("qcCREST",
       list.files(path = paste0("C:/Users", sep="/", Sys.info()[6], sep="/",
                                "DFO-MPO/PAC-SCA Stock Assessment (STAD) - Terminal CN Run Recon/Communal data" )),
       value=T)
  )
)


# Export new qcCREST Excel report --------------------------- 
# Will export to default Rmd working directory. 
writexl::write_xlsx(x=sheet_list,
           col_names=T, 
           path = 
             paste0(
               paste0(
                 "C:/Users", sep="/", Sys.info()[6], sep="/", 
                 "DFO-MPO/PAC-SCA Stock Assessment (STAD) - Terminal CN Run Recon/Communal data"
                 ),
               sep="/",
               paste0("qcCREST", sep="_", format(Sys.time(), sep="_", "%Y-%m-%d_%H%M"), ".xlsx")
             )
)


# Copy to SharePoint
file.copy(from = here("data_prep", "qcCREST.html"),
          to   = paste0(
                 "C:/Users", sep="/", Sys.info()[6], sep="/", 
                 "DFO-MPO/PAC-SCA Stock Assessment (STAD) - Terminal CN Run Recon/Communal data"
                 ))
```













